---
title: "Binary Logistic Regression and Modeling"
description: |
  Logistic Regression 
author:
  - name: Anna Talken
date: 02-08-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(GGally)
library(broom)
library(jtools)
library(kableExtra)
```


```{r, warning=FALSE}
#Read in the data
palmetto <- read_csv(here("_posts", "2021-02-08-logistic_regression", "palmetto.csv"),
                     col_types = cols(.default = 'c')) %>% 
  
  #convert the following variables to numeric class
  mutate(height = as.numeric(height)) %>% 
  mutate(width = as.numeric(width)) %>% 
  mutate(length = as.numeric(length)) %>% 
  mutate(green_lvs = as.numeric(green_lvs)) %>% 
  mutate(species = case_when(species == 1 ~ "Serenoa Repens", species == 2 ~ "Sabal Etonia")) %>% 
  
  #select for species, height, length, width, green leaves
select(species, height, length, width, green_lvs)
```

## Exploring differences in canopy height, length, width, and green leave abundance between two species of palmettos

```{r, fig.cap = "Figure 1. Exploratory scatterplot comparing the maximum canopy hieght and canopy width in Florida Palmettos between two species: Serenoa Repens and Sabal Etonia.", warning=FALSE}
  ggplot(data = palmetto, aes(x = width, y = height, color = species)) +
  geom_point(na.rm = TRUE) +
  scale_color_manual(values = c("cyan4", "coral")) +
  theme_minimal() +
  labs( title = "Comparison between canopy width and height in Florida Palmettos",
        x = "Width (cm)",
        y = "Height (cm)")
```

* There appears to be a positive correlation between canopy width and height within both species of Florida Palmettos. 


```{r, fig.cap = "Figure 2. A comparison of the number of green leaves found in each species: Serenoa Repens and Sabal Etonia", warning=FALSE}
# Second exploratory graph
ggplot(data = palmetto, aes(x = green_lvs, group = species, fill = species)) +
  geom_density(adjust = 1.5, alpha = 0.5) +
   theme_minimal() +
  labs( title = "Abundance of green leaves found within two species of Florida Palmettos",
        x = "Number of green leaves",
        y = "Density")
```
* The number of green leaves in Sabal Etonia that makes up the greatest density of total green leaves peaks around 3 to 4 leaves, whereas there is a more spread out density of green leaves for Serenoa Repens, with the highest density of green leaf numbers at around 6 to 7 leaves. 


```{r, fig.cap = "Figure 3. A comparison between the number of green leaves found and the widest length of the canopy within two species of Florida Palmettos: Serenoa Repens and Sabal Etonia.", warning=FALSE}
ggplot(data = palmetto, aes(x = green_lvs, y = length, color = species, shape = species)) +
  geom_point() +
  scale_color_brewer() +
  theme_minimal() +
   labs( title = "Comparison of the number of green leaves to canopy length",
        x = "Number of green leaves",
        y = "Length (cm)")
```
* There appears to be a positive correlation between number of green leaves and the widest length of the canopy for both species, though Serenoa Repens appears to have a braoder range of green leave abundance. 


## Binary logistic regression

```{r, warning=FALSE}
# Turning species into class 'factor'
palmettos <- palmetto %>% 
  mutate(species = as.factor(species))

#Creating a model for binomial logistic regression
palmetto_blr <- glm(species ~ height + length + width + green_lvs,
      data = palmettos,
      family = binomial)

# Get a tidy version of the model using broom
blr_tidy <- broom:: tidy(palmetto_blr)

#Creating a table output
blr_table <- kableExtra::kable(blr_tidy) %>% 
  kable_styling()

blr_table

```

## Model Predictions

```{r, warning=FALSE, message=FALSE}
#Converting log odds to the probability of prediting whether the plant is correctly for each observation

blr_fitted <- palmetto_blr %>% 
  broom::augment(type.predict = "response") %>%
  
  # creating a new column containing the species that the model predicts using a binary cutoff of 0.5 (less than 0.5 refers to Sabal Etonia with a factor of 0 and greater than 0.5 refers to Serenoa Repens, which has a factor of 1 in this model)
  mutate(species_new = case_when(.fitted < 0.5 ~ "Sabal Etonia", .fitted > 0.5 ~ "Serenoa Repens")) %>% 
  
  #If predicted species matches the original species in the dataset, it is considered correct
  mutate(model_classification = case_when(species_new == species ~ "Correct", species_new != species ~ "Incorrect"))
```

```{r, warning=FALSE, message=FALSE}
#creating table to include the species and how often the model correctly or incorrectly predicted the species

classified <- blr_fitted %>% 
  group_by(species, model_classification) %>% 
  summarise(Total = n()) %>% 
  pivot_wider(names_from = model_classification,
              values_from = Total) %>% 
  mutate(percent_correct = round(Correct/(Correct + Incorrect) * 100, 2))

kableExtra:: kable(classified,
                   caption = "Figure 1. The number of plants correctly and incorrectly predicted using a binary logistic regression model using two species: Sabal Etonia and Serenoa Repens. Percent correct indicates the percentage of correct predictions the model made when applied to the dataset on palmetto species in Florida from Abrahamson, W.G. 2019.",
      col.names = c("Species", "Correct", "Incorrect", "Percent Correct (%)")) %>% 
  kable_styling()
```
### Data Citation

Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5
